---
title: "DAL - Formative 1"
author: "B244545"
format:
  html:
    embed-resources: true
execute: 
  warning: false
---

**DUE ON THURSDAY 15 FEBRUARY AT NOON** (Week 5)

Complete the following tasks.

## Attach the tidyverse.

```{r}
#| label: setup

library(tidyverse)
```

## Import `data/coretta2018/token-measures.csv`

You can find information about the data here: <https://uoelel.github.io/dal/data/coretta2018/token-measures.html>.

```{r}
#| label: read-data

token_measures <- read_csv("data/coretta2018/token-measures.csv")
```

## Create the following plots.

For each, customise labels and add a description.

-   Scatter plot of `speech_rate` and `word_duration`. Set the alpha of the points to 0.3.

```{r}
#| label: Figure 1

token_measures |>
  ggplot(aes(speech_rate, word_duration)) +
  geom_point(alpha = 0.3) +
  labs(
    x = "Speech rate",
    y = "Word duration",
    title = "Scatter plot of speech rate and word duration",
  )

```

> Figure 1 shows the scatter plot of speech rate and word duration.

-   Scatter plot of `speech_rate` and `v1_duration`, coloured by `c2_phonation`. Set the alpha of the points to 0.5. Use the turbo viridis palette.

```{r}
#| label: Figure 2

token_measures |>
  ggplot(aes(speech_rate, v1_duration, colour = c2_phonation)) +
  geom_point(alpha = 0.5) +
  scale_color_viridis_d(option = "H") +
  labs(
    x = "Speech rate",
    y = "V1 duration",
    colour = "C2 phonation",
    title = "Scatter plot of speech rate and v1 duration coloured by c2 phonation",
  )

```

> Figure 2 shows the scatter plot of speech rate and v1 duration, by c2 phonation.

-   Scatter plot of `f2` and `f1`, coloured by `vowel`. Set the alpha of the points to 0.5. Use the Dark2 brewer palette.


```{r}
#| label: Figure 3

token_measures |>
  ggplot(aes(f2, f1, colour = vowel)) +
  geom_point(alpha = 0.5) +
  scale_colour_brewer(palette = "Dark2") +
  labs(
    x = "f2 (HZ)",
    y = "f1 (Hz)",
    colour = "Vowel",
    title = "Scatter plot of f2 and f1 coloured by vowel",
  )

```

> Figure 3 shows the scatter plot of f1 and f2, by vowels.
